### 2014-05-27 Paper1/paper1.tex
Same
Updated to reflect weekend edits

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from force-disturbed arm movement, examine its parameter sensitivity, and validate it on human reaching data. We demonstrate that inverting popular arm modeling approaches algebraically allows determination of movement intent from recordings of force disturbances and arm movement. We then extend this to a more general approach reveals strengths, limitations, and underlying assumptions. After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching. We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching. Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints. The robustness of this approach indicated that it was suitable for determining reaching intent from human point-to-point reaching movements disturbed by large force pulses. Intended trajectory showed no change from undisturbed reaching for 150 ms after the onset of disturbance. Beyond this point, stiffening of the arm, change in intent, or both usually lead to deviation from undisturbed reaching. Not all disturbed reaches showed evidence of change in intent. Unlike $\lambda$ models, the reaching intent determined by our model is physiologically plausible. This success allows real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals. In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent. While not explicitly examined, this approach is also valid for determination intent is more complex situations like team sports and national security.

@@

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. More complex models can be used without any loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripetal effects, and $E$ is externally-applied torque that the system might interact with. Dots ($\dot{q}, \ddot{q}$) are used to denote time derivatives of the joint angles. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
Where a hat ($\hat{ }$) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ signifies the unknown desired trajectory that we seek to discover.

@@

Note $x_e$ vanishes, recovering our familiar model. While a proper choice of $x_e$ is critical to control, our approach does not require a model relating equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}.

In summary, if the highest order coefficient of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory. This general form reveals the conditions needed to solve for the moving intent of an arbitrary dynamic process.

@@

\subsection*{Human Subjects}
The human data trajectories analyzed here are drawn from a four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board. Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

@@

As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. As demonstrated in figure 4, the method passes this test until 150 milliseconds after the onset of disturbance. Further

\section*{Discussion}
We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances. Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching. Significant misestimation could occur if the stiffness was inaccurately modelled. We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target until 120-150 ms before responding. This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.

@@

This approach explains the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$. By deriving the technique in a general form, we discovered that our $x_d$ can be used to determine the muscle equilibrium of a $\lambda$ model and that they are only equivalent when $x_d$ is not changing.  Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass. This provides a striking and falsifiable prediction: reaching intention cannot jump abruptly.

@@

Determination of intent may facilitate may facilitate motor training and stroke recovery. Error augmentation, which presently relies on dictating the reaching intent, has demonstrated the capacity to increase and speed up learning in healthy patients \cite{patton2004robot} and following stroke \cite{patton2006evaluation}. This static, dictated intent could be replaced By instead scaling the magnitude of the difference between the desired and realized trajectory, error augmentation during undirected reaching and exploration becomes possible. This permits both error reduction and error magnification. Partial error cancellation would allow learning to take place without harm to task goals. As demonstrated by the success of the challenge-point framework \cite{guadagnoli2004challenge}, dynamic variation of augmentation as task learning progresses can be beneficial. With this extraction, error can be measured and augmented in real-time even without an explicit task, potentially enabling wider utility and application.

Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible. It is important to reiterate that because the human arm’s mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory. By using tests such as the comparison with undisturbed straight line reaching employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.  Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting. Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions? They do. Similarly, compositional claims, such as whether movements are comprised of subunits, should now be directly testable. Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit hypothesized cost functions.

The plant model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance. The plant (such as the actions of a crowd in a crowded square, the motions of the opposing team,) may be too complex to model with simple linear transformations as we present above. However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed. It is not necessarily to allow a bomb to explode or the opposing team to be victorious to understand the plan behind their action.

@@

The approximation of stiffness appears to be the most critical factor in accuracy in estimating intent. While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach. Even as the model is insensitive to normal variations, a human can volitionally halve or double their stiffness \cite{franklin2007endpoint}. Co-contraction is plausible and has significant impact on accuracy. These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations. Additionally, real-time determination of intent may allow real-time determination of stiffness.

Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting. More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result. Our results point to the possibility of the impedance changing during motion, perhaps as a function of time. As mentioned above, while this may be perceived as a shortcoming, it was necessary and potentially facilitates real-time determination of stiffness in future work.

The general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process.

@@

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig1.eps}
\end{center}
\caption{
{\bf Simulated data illustrating tautology of extraction absent parameter error across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial. Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory. Extraction to recover intention from arm and force trajectory follows.
}
\label{fig:synthetic}
\end{figure}

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig2.eps}
\end{center}
\caption{
{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Parameter values are varied in a Sobol-distributed fashion according to Table 1. First order sensitivity, in purple, is the error that would be removed if the parameter was fixed at its nominal value. Total sensitivity, in teal, shows error that would remain if all other parameters were fixed at their nominal values.
}
\label{fig:sensitivity}
\end{figure}

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig3.eps}
\end{center}
\caption{
{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the green baseline, as a force pulse, gray arrows, are applied. This subject's extracted desired trajectories, in red (baseline in faded red), do not significantly differ from baseline during force application, but show some reaction afterwards.
}
\label{fig:anecdotes}
\end{figure}

\begin{figure}[!ht]
\begin{center}
%\includegraphics[width=4in]{fig4.eps}
\end{center}
\caption{
{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) begin to deviate from undisturbed hand errors (green dots, band) beginning 95 ms after the onset of disturbance. Movement intention (red dots, band) deviates at 150ms. The hand rejoins the intention approximately 500ms after the onset of disturbance.
}
\label{fig:grouptrends}
\end{figure}

@@

\begin{table}[!ht]
\caption{
\bf{Arm Parameter Values and Sensitivities}}
\begin{tabular}{|c|c|c c|c c|}
\hline
Parameter Name &
Units &
Nominal &
SD &
$S$ &
$S_T$ \\ \hline
Upper Arm Length ($L_1$) &
m &
0.33 &
0.01 &
0 &
0.10 \\
Forearm Length ($L_2$) &
m &
0.34 &
0.01 &
0 &
0.03 \\
Upper Arm Center of Mass Ratio ($\frac{L_{m1}}{L_1}$) &
1 &
0.436 &
0.0695 &
0 &
0.02 \\
Forearm Center of Mass Ratio ($\frac{L_{m2}}{L_2}$) &
1 &
0.682 &
0.0431 &
0 &
0.02 \\
Gross Body Mass ($m_g$) &
kg &
79.36 &
3.1 &
0 &
0.02 \\
Upper Arm Mass Ratio ($\frac{m_1}{m_g}$) &
1 &
0.028 &
0.0029 &
0 &
0.01 \\
Forearm Mass Ratio ($\frac{m_2}{m_g}$) &
1 &
0.022 &
0.0025 &
0 &
0.06 \\
Upper Arm Radius of Gyration Ratio &
1 &
0.322 &
0.0161 &
0 &
0.01 \\
Forearm Radius of Gyration Ratio &
1 &
0.468 &
0.0234 &
0 &
0.01 \\
Shoulder Parallel Coordinate &
m &
-0.057 &
0.02 &
0.02 &
0.18 \\
Shoulder Perpendicular Coordinate &
m &
0.88 &
0.02 &
0 &
0.08 \\
Force Sensor Miscalibration $x\text{-axis}$ &
N &
0 &
0.231 &
0 &
0.08 \\
Force Sensor Miscalibration $y\text{-axis}$ &
N &
0 &
0.1067 &
0.03 &
0.08 \\
Force Sensor Gaussian Noise SD $x\text{-axis}$ &
N &
0 &
0.1653 &
0 &
0.01 \\
Force Sensor Gaussian Noise SD $y\text{-axis}$ &
N &
0 &
0.2869 &
0 &
0.01 \\
Torque-Invariant Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.06 &
0.31 \\
Torque-Varying Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.0 &
0.15 \\
Damping-to-Stiffness Ratio ($k_d$) &
$\text{sec}^{-1}$ &
0.0833 &
0.0125 &
0 &
0.08 \\
Reflex Impedance Scale Factor &
1 &
0.02 &
0.003 &
0 &
0.04 \\
Reflex Damping to Stiffness Ratio ($g_d$) &
$\text{sec}^{-1}$ &
2 &
0.3 &
0 &
0.04 \\ \hline
\end{tabular}
\begin{flushleft}Synthetic model parameters and their associated mean (nominal) values and standard deviations (SD) used to determine the sensitivity indices of Saltelli et al. \cite{saltelli2010variance}. Also shown are the resulting sensitivity indices. Sensitivity indices less than one-thousandth of the total sensitivity were reported as zero. Total sensitivity was $1.3 \mu m^2$, showing very small average deviations when using this approach. First order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$ can be interpreted as the fraction left after correcting all other factors.
\end{flushleft}
\label{tab:parameters}
\end{table}

>>>>

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from force-disturbed arm movement, examine its parameter sensitivity, and validate it on human reaching data. We demonstrate that inverting popular arm modeling approaches algebraically allows determination of movement intent from recordings of force disturbances and arm movement. We then extend this to a more general approach, which reveals strengths, limitations, and underlying assumptions. After populating our model with measured lengths and masses as well as impedance values from the literature, we perform sensitivity analysis to determine the accuracy of the model relative to the uncertainty naturally inherent in point-to-point reaching. We find that sensitivity to parameter inaccuracy is on the order of microns and lower than the inaccuracy of human reaching. Further, most of the sensitivity results from uncertainty with regard to the stiffness of the joints. The robustness of this approach indicated that it was suitable for determining reaching intent from human point-to-point reaching movements disturbed by large force pulses. Intended trajectory showed no change from undisturbed reaching for 150 ms after the onset of disturbance. Beyond this point, stiffening of the arm, change in intent, or both usually lead to deviation from undisturbed reaching. Not all disturbed reaches showed evidence of change in intent. Unlike $\lambda$ models, the reaching intent determined by our model is physiologically plausible. This success allows real-time determination of error that can free learning enhancement and treatment methods like error augmentation and challenge point from needing to prescribe explicit task goals. In addition, extraction from recorded human motion enables direct inspection of and hypothesis testing on the contents of human motor intent. While not explicitly examined, this approach is also valid for determination intent is more complex situations like team sports and national security.

@@

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. More complex models can be used without any loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripetal effects, and $E$ is externally-applied torque that the system might interact with. Dots ($\dot{q}, \ddot{q}$) are used to denote time derivatives of the joint angles. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
Where a hat ( $\hat{ }$ ) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ signifies the unknown desired trajectory that we seek to discover.

@@

Note $x_e$ vanishes, recovering our familiar model. While a proper choice of $x_e$ is critical to control, our approach does not require a model relating equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.

While we treat both the process and actuator as dynamic models in our derivation, both the derivation and the technique are agnostic to the process model. Instead of a dynamic model, it could be a lookup table. For instance, a light switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined. This allows determination of intention in many situations of interest, even where the process is irreducibly complex.

@@

\subsection*{Human Subjects}
The human data trajectories analyzed here are drawn from four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board. Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

@@

As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. As demonstrated in figure 4, the method passes this test until 150 milliseconds after the onset of disturbance.

\section*{Discussion}
We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances. Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching. Significant misestimation could occur if the stiffness was inaccurately modeled. We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target until 120-150 ms before responding. This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.

@@

This modeling approach explains the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$. By deriving the technique in a general form, we discovered that our $x_d$ should be equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model. $x_d$ and $\lambda$ are only equivalent when $x_d$ is not changing.  Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass. This provides a striking and falsifiable prediction: reaching intention cannot jump abruptly.

@@

Determination of intent may facilitate motor training and stroke recovery. Error augmentation, which presently relies on dictating the reaching intent, has demonstrated the capacity to increase and speed up learning in healthy patients \cite{patton2004robot} and following stroke \cite{patton2006evaluation}. Augmentation of difference from static, dictated intent could be replaced with scaling of the magnitude of the difference between the desired and realized trajectory. This makes error augmentation during undirected reaching and exploration possible, including both error reduction and error magnification. Partial error cancellation would allow learning to take place without harm to task goals. As demonstrated by the success of the challenge-point framework \cite{guadagnoli2004challenge}, dynamic variation of augmentation as task learning progresses can be beneficial. With this extraction, error can be measured and augmented in real-time even without an explicit task, potentially enabling wider utility and application.

Determination of intended trajectories from force-disturbed movements allows hypothesis testing that was not previously possible. It is important to reiterate that because the human arm's mass is almost always invertible, this approach can take \textbf{any} ``candidate model'' of the human arm and create a filter that estimates the intended trajectory. By using tests such as the comparison with undisturbed straight line reaching employed in this paper, one can evaluate which candidate produces the most plausible intended trajectory, supporting one model over another.  Instead of comparing two generative models to see which better approximates movement, it is now possible to perform element-by-element fitting. Mid-movement replanning is now falsifiable: do intended trajectories from disturbed motions differ from undisturbed motions? They do. Similarly, compositional claims, such as whether movements are comprised of subunits, should now be directly testable. Finally, optimal feedback control models can now gain direct access to an intended trajectory to better understand and fit hypothesized cost functions.

The process model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance. The plant (such as the actions of a crowd in a crowded square or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above. However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed. It is not necessary to allow a bomber to succeed or an opposing team to be victorious to understand the plan behind their action.

@@

The approximation of stiffness appears to be the most critical factor to accuracy in estimating intent. While serious trial-to-trial stiffness misestimation is unlikely, co-contraction in response to disturbance might plausibly double stiffness mid-reach. Even as the model is insensitive to normal variations, a human can halve or double their stiffness \cite{franklin2007endpoint} significantly reducing determination accuracy. These limitations are readily overcome by performing this sensitivity analysis on each force environment of interest and taking measures to detect co-contraction and hence stiffness alterations. Additionally, real-time determination of intent may allow real-time determination of stiffness.

Our simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting. More accurate approaches would provide us with too many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result. Our results point to the possibility of the impedance changing during motion, perhaps as a function of time. As mentioned above, while this may be perceived as a shortcoming, it was necessary and can potentially be overcome in future work.

The general intent extraction approach presented here reveals the conditions needed to solve for the moving intent in an arbitrary dynamic process. Sensitivity analysis for this approach applied to the human arm demonstrated favorable conditions for determination of intent during force-disturbed reaching. Reaching intent became significantly different from baseline reaching only after enough time had elapsed for the disturbance to be processed and descending motor commands to change. This window into intent should allow advances in arm modeling, motor training, and human-machine interaction.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{fig1.eps}
\end{center}
\caption{
{\bf Simulated data illustrating tautology of extraction absent parameter error across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial. Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory. Extraction to recover intention from arm and force trajectory follows.
}
\label{fig:synthetic}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{fig2.eps}
\end{center}
\caption{
{\bf Model sensitivity testing (from the variance-based method of Saltelli et al.\cite{saltelli2010variance}) reveals that the model is mostly sensitive to stiffness.} Parameter values are varied in a Sobol-distributed fashion according to Table 1. First order sensitivity, in purple, is the error that would be removed if the parameter was fixed at its nominal value. Total sensitivity, in teal, shows error that would remain if all other parameters were fixed at their nominal values.
}
\label{fig:sensitivity}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{fig3.eps}
\end{center}
\caption{
{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the green baseline, as a force pulse, gray arrows, are applied. This subject's extracted desired trajectories, in red (baseline in faded red), do not significantly differ from baseline during force application, but show some reaction afterwards.
}
\label{fig:anecdotes}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{fig4.eps}
\end{center}
\caption{
{\bf Comparison of subjects' perpendicular errors across disturbed and undisturbed conditions reveals that intention is statistically indistinguishable from undisturbed reaching up until 100ms after the onset of movement.} Hand errors in the presence of disturbance (black dots, black band is median $\pm$ simultaneous 95\% confidence interval on the median) begin to deviate from undisturbed hand errors (green dots, band) beginning 95 ms after the onset of disturbance. Movement intention (red dots, band) deviates at 150ms. The hand rejoins the intention approximately 500ms after the onset of disturbance.
}
\label{fig:grouptrends}
\end{figure}

@@

\begin{table}[!ht]
\caption{
\bf{Arm Parameter Values and Sensitivities}}
\begin{tabular}{|c|c|c c|c c|}
\hline
Parameter Name &
Units &
Nominal &
SD &
$S$ &
$S_T$ \\ \hline
Upper Arm Length ($L_1$) &
m &
0.33 &
0.01 &
0 &
0.10 \\
Forearm Length ($L_2$) &
m &
0.34 &
0.01 &
0 &
0.03 \\
Upper Arm Center of Mass Ratio ($\frac{L_{m1}}{L_1}$) &
1 &
0.436 &
0.0695 &
0 &
0.02 \\
Forearm Center of Mass Ratio ($\frac{L_{m2}}{L_2}$) &
1 &
0.682 &
0.0431 &
0 &
0.02 \\
Gross Body Mass ($m_g$) &
kg &
79.36 &
3.1 &
0 &
0.02 \\
Upper Arm Mass Ratio ($\frac{m_1}{m_g}$) &
1 &
0.028 &
0.0029 &
0 &
0.01 \\
Forearm Mass Ratio ($\frac{m_2}{m_g}$) &
1 &
0.022 &
0.0025 &
0 &
0.06 \\
Upper Arm Radius of Gyration Ratio &
1 &
0.322 &
0.0161 &
0 &
0.01 \\
Forearm Radius of Gyration Ratio &
1 &
0.468 &
0.0234 &
0 &
0.01 \\
Shoulder Parallel Coordinate &
m &
-0.057 &
0.02 &
0.02 &
0.18 \\
Shoulder Perpendicular Coordinate &
m &
0.88 &
0.02 &
0 &
0.08 \\
Force Sensor Miscalibration $x\text{-axis}$ &
N &
0 &
0.231 &
0 &
0.08 \\
Force Sensor Miscalibration $y\text{-axis}$ &
N &
0 &
0.1067 &
0.03 &
0.08 \\
Force Sensor Gaussian Noise SD $x\text{-axis}$ &
N &
0 &
0.1653 &
0 &
0.01 \\
Force Sensor Gaussian Noise SD $y\text{-axis}$ &
N &
0 &
0.2869 &
0 &
0.01 \\
Torque-Invariant Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.06 &
0.31 \\
Torque-Varying Impedance Misestimation Ratio &
1 &
1 &
0.15 &
0.0 &
0.15 \\
Damping-to-Stiffness Ratio ($k_d$) &
$\text{sec}^{-1}$ &
0.0833 &
0.0125 &
0 &
0.08 \\
Reflex Impedance Scale Factor &
1 &
0.02 &
0.003 &
0 &
0.04 \\
Reflex Damping to Stiffness Ratio ($g_d$) &
$\text{sec}^{-1}$ &
2 &
0.3 &
0 &
0.04 \\ \hline
\end{tabular}
\begin{flushleft}Synthetic model parameters and their associated mean (nominal) values and standard deviations (SD) used to determine the sensitivity indices of Saltelli et al. \cite{saltelli2010variance}. Also shown are the resulting sensitivity indices. Sensitivity indices less than one-thousandth were reported as zero. Total variance was $1.3 \mu m^2$, showing very small average deviations when using this approach. First-order sensitivity, $S$, can be interpreted as fraction of variance removable by perfectly correcting a factor while total sensitivity, $S_T$, can be interpreted as the fraction left after correcting all other factors.
\end{flushleft}
\label{tab:parameters}
\end{table}
